STAT 672 Final Project: Stochastic Gradient Descent

### 2018-05-04 paper/paper.tex
final

@@

\title{STAT 672 final project:\\ Stochastic gradient descent}

@@

\subsection{Parametric vs. non-parametric}

@@

\begin{table}[t]
	\centering
	\caption{Asymptotic comparison of GD and SGD}
	\begin{tabular}{|l l l|}
		\hline
		& \textbf{GD} & \textbf{SGD} \\
		Iterations to accuracy $\rho$ & $\log \frac{1}{\rho}$ & $\frac{1}{\rho}$ \\
		Time per iteration & $n$ & 1 \\
		Time to accuracy $\rho$ & $n \log \frac{1}{p}$ & $\frac{1}{\rho}$ \\
		\hline
	\end{tabular}
\end{table}

@@

Averaged SGD (ASGD)---first proposed in 
\cite{polyak1992acceleration}---is another variation. Typically, we consider
$\hat{w}$ to be the final iteration
of SGD, when some stop condition is met. For example, in a situation where we
run $N$ iterations, $\hat{w} = w^{(N)}$. In ASGD, we consider $\hat{w}$ 
the average across all iterations.

@@

@@

If a Silicon Valley press release uses words like ``artificial intelligence'',
``machine learning'', and ``big data'' to describe a new product or service, it very likely
uses SGD in some way. Such is the extent of SGD's popularity. 
Below are some prominent examples of SGD's practical
applications.

>>>>

\usepackage{geometry}
\geometry{margin=1.5in}

@@

\title{STAT 672 Final Project: Stochastic Gradient Descent}

@@

\subsection{Parametric vs. non-parametric approaches}

@@

\begin{table}[h]
	\centering
	\caption{Asymptotic comparison of GD and SGD}
	\begin{tabular}{|l l l|}
		\hline
		& \textbf{GD} & \textbf{SGD} \\
		Iterations to accuracy $\rho$ & $\log \frac{1}{\rho}$ & $\frac{1}{\rho}$ \\
		Time per iteration & $n$ & 1 \\
		Time to accuracy $\rho$ & $n \log \frac{1}{p}$ & $\frac{1}{\rho}$ \\
		\hline
	\end{tabular}
\end{table}

@@

Averaged SGD (ASGD)---proposed in
\cite{polyak1992acceleration}---is another variation. Typically, we consider
$\hat{w}$ to be the final iteration
of SGD, when some stop condition is met. For example, in a situation where we
run $N$ iterations, $\hat{w} = w^{(N)}$. In ASGD, we consider $\hat{w}$ 
the average across all iterations.

@@

\subsection{Neural networks and back-propagation}

Neural networks are another class of statistical learning
models for which SGD plays a fundamental role in estimating weights, in
particular through a process called back-propagation. Space
constraints disallow a detailed examination along the lines of that conducted
for SVM. This subsection aims only to note that SGD is widely used with neural
networks, and that this usage provides the practical motivation behind many of
the extensions to the SGD algorithm covered in Section 2.3. This is not to say
that these techniques are \textit{solely} intended for neural networks---indeed,
many of them precede neural networks' current popularity---but
their utility for neural networks is part of why the relevant papers have
thousands of citations.

Neural networks
often use non-convex loss functions that have multiple local minima in
which SGD can get ``lost'' while searching for global minimum.
Acceleration and momentum (2.3.2) thus are required to achieve effective
training, particularly on ``deep'' networks with many hidden layers
(\cite{sutskever2013importance}). Neural networks are
often extremely complex and high-dimensional and hence require enormous training
data (\cite{lecun2015deep}); hence, speeding up computation via parallelization
(2.3.5) or GPU computing is common.

@@

If a Silicon Valley press release uses phrases like ``artificial intelligence'',
``machine learning'', and ``big data'' to describe a new product or service, it very likely
relies upon SGD in some way. Such is the extent of SGD's popularity. 
Below are some prominent examples of SGD's practical
applications.
